Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-08-12 writeup/gz_hubble_data.tex
first draft of reorganizing section 4

@@

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
surface brightness ($\mu$; intrinsic) and redshift ($z$;
extrinsic).

We model the debiased vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy as:

\begin{equation}
f_{\mu,z} = \left(f_{\mu,z=0.3}\right) \times e^{{\frac{z-z_0}{\zeta}}},
\label{eqn:fzeta}
\end{equation}

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction, the simple exponential form of
Equation~\ref{eqn:fzeta} is modified to bound the debiased vote fractions
between $f$ and 1:

\begin{equation}
f_\textrm{features,debiased} = 1 - (1 - f_\mathrm{features,weighted})e^{\frac{z-z_0}{\hat\zeta}}
\label{eqn:fzeta_mod}
\end{equation}

\noindent where $f_\mathrm{features,weighted}$ is the weighted vote fraction
described in Section~\ref{ssec:weighting}.

\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the ``debiased''
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the previous section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

The unshaded regions in Figure~\ref{fig:f_vs_f} define discrete ranges of redshift,
surface brightness, and \ffeatures{} within which a galaxy must lie in order for the
debiased correction to be confidently applied.
While the appropriate correctable regions were defined as discrete bins, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimised. The resulting hulls define the
correctable and lower-limit regions for categorising the \hst{} galaxies. The
results of this method and final categorisation of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be debiased
since the relationship between \ffeaturesz{} and \ffeaturesrest{} is not
monotonic, and 56\% cannot be debiased since they either have an unknown
redshift or insufficient numbers of \ferengi{} images in their $z$-$\mu$ bin
to determine a reliable value of $\zeta$.

>>>>

\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}
The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
surface brightness ($\mu$; intrinsic) and redshift ($z$;
extrinsic).

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the ``debiased''
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the following section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

The unshaded regions in Figure~\ref{fig:f_vs_f} define discrete ranges of redshift,
surface brightness, and \ffeatures{} within which a galaxy must lie in order for the
debiased correction to be confidently applied.
While the appropriate correctable regions were defined as discrete bins, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimised. The resulting hulls define the
correctable and lower-limit regions for categorising the \hst{} galaxies. The
results of this method and final categorisation of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be debiased
since the relationship between \ffeaturesz{} and \ffeaturesrest{} is not
monotonic, and 56\% cannot be debiased since they either have an unknown
redshift or insufficient numbers of \ferengi{} images in their $z$-$\mu$ bin
to determine a reliable correction term ($\zeta$, described in section ~\ref{ssec:zeta}).

@@

For the ``correctable'' sample of galaxies, we observe a decline in the vote fraction
\ffeatures{} with increasing simulated redshift for each unique galaxy. 
We model this relationship with a simple exponential
function which bounds $f_{\mu, z>0.3}$  between $f_{\mu, z=0.3}$ and 0:

\begin{equation}
f_{\mu,z} = 1 - (1 - f_{\mu,z=0.3})e^{\frac{z-z_0}{\hat\zeta}}
\label{eqn:fzeta}
\end{equation}

%old equation
%We model the debiased vote fraction $f_{\mu,z}$ for an
%artificially-redshifted galaxy as:

%\begin{equation}
%f_{\mu,z} = \left(f_{\mu,z=0.3}\right) \times e^{{\frac{z-z_0}{\zeta}}},
%\label{eqn:fzeta}
%\end{equation}

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction, we use Equation~\ref{eqn:fzeta}, rearranged for clarity:

\begin{equation}
f_\textrm{features,debiased} = 1 - (1 - f_\mathrm{features,weighted})e^{\frac{-(z-z_0)}{\hat\zeta}}
\label{eqn:fzeta_mod}
\end{equation}

\noindent where $f_\mathrm{features,weighted}$ is the weighted vote fraction
described in Section~\ref{ssec:weighting}, and here $f_{\rm features,debiased}$ is bounded 
between $f_{\rm features,weighted}$ and 1.
